---
output: 
  github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  dpi=600,
  fig.width=7)
```

# Computing Activation Zone Boundaries for [SOTA](https://www.sota.org.uk/) Summits in W7W

<!-- badges: start -->

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/benmarwick/computing-sota-az-boundaries/master?urlpath=rstudio)

<!-- badges: end -->


### Background

The goal of this repository is to generate polygons that indicate the [activation zones](https://www.sota.org.uk/Blog/2017/07/08/In-The-Zone) for [SOTA](https://www.sota.org.uk/) summits in the W7W area (Washington State, USA). Here is how the [SOTA FAQ](https://www.sota.org.uk/Joining-In/FAQs) defines the activation zone:

> Q. What is the Activation Zone?

> A. Often it is inconvenient to operate from the actual highest point of a summit, there may be structures there, or on a frequently visited summit there may be too many people about. The SOTA rules permit the operating position to be away from the summit but not more than 25 vertical metres below the summit. If you draw a contour line on a map 25 metres below the summit, the area within this contour is the Activation Zone. See this [MT blog posting](https://www.sota.org.uk/Blog/2017/07/08/In-The-Zone) for more information. If you operate from outside the AZ, the activation is not valid and will score no points.

### Prior work on computing SOTA activation zones 

- the web app https://activation.zone/ by [Ara N6ARA](https://n6ara.com/) which uses 30m elevation data with global coverage provided by the [elevation Python package](https://pypi.org/project/elevation/). This is fast and easy to use, but relatively low resolution. 
- On [sotl.as](https://sotl.as/) activation zones are visible in the map view for HB/HB0 (calculated using [swissALTI3D](https://www.swisstopo.admin.ch/de/hoehenmodell-swissalti3d) data from swisstopo (spatial resolution 0.5 m, accuracy ± 0.3 – 3 m (1σ) depending on the region) and OE (calculated using BEV ALS DTM data (spatial resolution 1 m, accuracy generally ± 0.5 m, may vary in high altitude). The OE activation zones were computed by [Tobi OE7TOK](https://reflector.sota.org.uk/t/activation-zones-for-oe-on-sotlas/34629). 
- [Caltopo](https://caltopo.com/) can also be used to compute activation zones using the DEM shading tool on a desktop computer or smartphone (see [Tim N7KOM's](https://www.etsy.com/shop/N7KOMPortableRadio) [tutorial](https://www.youtube.com/watch?v=UixA1Fc4D1c) on how to do this). CalTopo's elevation data is up to 1 meter resolution in many areas, which is based on LIDAR scans from the USGS's 3DEP program. This is a high resolution freely available tool for on-the-fly computation of activation zone areas. However, it can be a bit fiddly to use when using on a small screen on a summit.  

### This project

```{r}
#| echo: false

library(tidyverse)

n_activation_zones <- 
  length(list.files(pattern = "[0-9]{3}.geojson", recursive = TRUE))

# how many summits in the SOTA program? Get data from SOTA API
library(rvest)
summits_in_sota <-  
  httr::GET("https://api2.sota.org.uk/api/associations/w7w") %>% 
            httr::content(., as = "text") %>% 
            jsonlite::fromJSON() %>% 
            purrr::pluck("regions") %>% 
            tibble() %>% 
  select(regionName, 
         regionCode,
        summits )

n_total_summits_w7w <- sum(summits_in_sota$summits)

perc <- round(n_activation_zones / n_total_summits_w7w * 100, 1)
```


The goal of these project is generate activation zone polygons as GeoJSON files that can be integrated into the [sotl.as map](https://sotl.as/summits/W7W) for W7W. This is inspired by the work done for HB/HB0 and OE, where operators can view the activation zone polygon directly in the sotl.as map. Currently this project has computed polygons for activation zones for `r n_activation_zones` summits in the W7W area. The [official SOTA database](https://www.sotadata.org.uk/en/association/W7W) has `r n_total_summits_w7w` summits recorded for W7W, so `r perc`% of all summits in W7W have an activation zone. We're unable to compute activation zones for summits in areas that currently do not have lidar imagery.

Here I use the R programming language to access the [Washington State's Department of Natural Resources public LIDAR portal](https://lidarportal.dnr.wa.gov/). In brief, my method was to automatically download a small raster file of elevation data for each summit, compute the activation zone polygon, and save it to a GeoJSON file. Please inspect the R Markdown files in this repository for more details of the calculations. I visually inspected each activation zone polygon to ensure there were no errors, but if you find something that looks odd, please let me know and I'll take another look.  

Currently this repository has GeoJSON files for activation zones in the following regions in the W7W association:

```{r}
#| echo: false

# how many summits do we have activation zones for?
w7w_regions <- 
list.files(pattern = "W7W") %>% 
  str_remove_all("W7W_")

summits_per_region_with_az <- 
  map_int(w7w_regions,
    ~list.files(pattern = paste0(.x, ".*geojson"),
                recursive = TRUE) %>% 
      length()) %>% 
  tibble(summits_with_activation_zones = .,
         regionCode = w7w_regions)

# combine the tables
summit_counts_and_per_with_az <- 
summits_in_sota %>% 
  left_join(summits_per_region_with_az,
            join_by("regionCode")) %>% 
  mutate(percent_summits_with_az = round(summits_with_activation_zones / summits * 100, 1 )) %>% 
  rename(Region = regionName ,
        Code = regionCode,
        `Number of summits` = summits,
       `Number of activation zones` =  summits_with_activation_zones,
       `Percentage of summits\n with activation zones` = percent_summits_with_az) %>% 
  arrange(desc(`Percentage of summits\n with activation zones`),
          desc( `Number of activation zones`) )

knitr::kable(summit_counts_and_per_with_az, format = 'html')

```



### Licenses

**Text and figures :** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** see [LICENSE.md](LICENSE.md)

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/)

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.


