---
output: github_document
always_allow_html: true
leafletmap: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  dpi=600,
  fig.width=7)
```

# Computing Activation Zone Boundaries for [SOTA](https://www.sota.org.uk/) Summits in W7W-OK 

Okenagen region (OK)


### Prepare the summit locations and bounding boxes

Let's load the libraries we'll need

```{r}
library(raster)
library(terra)
library(tidyterra)
library(ggplot2)
library(sf)
library(tidyverse)
library(ggrepel)
library(ggspatial)
library(httr2)
```

Import a GeoJSON file containing the locations and elevations of SOTA summits. I got these data from the SOTA API.

```{r}
# get all summits via the SOTA API

library(rvest)
gjsf <-  httr::GET("https://api2.sota.org.uk/api/regions/W7W/FR") %>% 
            httr::content(., as = "text")%>% 
            jsonlite::fromJSON() %>% 
            purrr::pluck("summits") %>% 
            tibble() %>% 
  st_as_sf(coords =  c("gridRef1",
                       "gridRef2"),
           crs = st_crs(4326))

this_dir <- here::here("W7W_FR")
```

Extract the location coordinates into columns so we can use them later, and create a column for elevation in meters because our raster data is in meters.

```{r}
# make sf data frame and get coords into cols we can use
gjsf_elev <- 
gjsf %>% 
  mutate(elev_m = altM,
         elev_ft = altFt) %>% 
  mutate(x = st_coordinates(geometry)[,"X"],
         y = st_coordinates(geometry)[,"Y"]) 
```

Create a square buffer area around each summit, with the summit in the center. This is the area that we will download a raster of elevation data to compute the activation zone. This area is good enough for most summits, but for a few summits with very large activation zones it's too small, we must increase the buffer side length to ensure we capture the full area of the activation zone. I did this manually for the small number of summits that needed it. 

```{r}
# make a single square buffer for each point
# Function to create the square buffers
# https://stackoverflow.com/a/70372149/1036500
bSquare <- function(x, a) {
  a <- sqrt(a)/2
  return( sf::st_buffer(x, dist = a, nQuadSegs=1, 
                        endCapStyle = "SQUARE") )
}

# get a buffer zone of a certain area
buffer_side_length <- 1e7 # 100 is a 10x10, this is the value we manually adjust to manage buffer area, to capture the biggest AZ
gjsf_elev_buf <- 
  bSquare(gjsf_elev,
          buffer_side_length)

gjsf_elev_buf_sq <- vector("list", nrow(gjsf_elev_buf))
for(i in 1:nrow(gjsf_elev_buf)){
  gjsf_elev_buf_sq[[i]] <- 
    st_bbox(gjsf_elev_buf[i,]) %>%
    st_as_sfc() %>%
    st_as_sf()
}

gjsf_elev_buf_sq_df <- 
  bind_rows(gjsf_elev_buf_sq)
```

Take a look at the buffer zones and summit points altogether

```{r}
ggplot() + 
  geom_sf(data = gjsf_elev_buf_sq_df,
           colour = "red", fill = NA) +
  geom_sf(data = gjsf,
          size = 0.1) +
  coord_sf() +
  annotation_scale(location = "bl", 
                   width_hint = 0.5) +
  theme_minimal()
```

Convert the buffer zone into a list of bounding box coordinates to input into the raster download function

```{r}
# get list of bounding boxes coords
gjsf_elev_buf_sq_bbx <- vector("list", nrow(gjsf_elev_buf_sq_df))
for(i in 1:nrow(gjsf_elev_buf_sq_df)){
  gjsf_elev_buf_sq_bbx[[i]] <- 
    gjsf_elev_buf_sq_df$x[i] %>% 
    st_coordinates() %>% 
    as.data.frame() %>% 
    select(X, Y)
}
```

### Iterate over each summit to compute the activation zone polygon

Here is a loop that takes each summit and its bounding box and:

-   downloads some raster tiles to cover the bounding box area
-   if there are multiple rasters for an area, merge them, and if they have different resolutions, downsample the high resolution rasters so they all have the same resolution 
-   then selects all the pixels in the raster that have an elevation value between the summit elevation and 25m below that elevation
-   then draw a polygon around those pixels
-   then drop polygons that don't contain the summit point, or are with 20 m of it (since a few summits are outside of their activation zone)
-   then write the polygon of the activation zone to a GeoJSON file

This chunk of code takes several hours to run and downloads raster tiles on each iteration of the loop. 


```{r, results='hide', eval=FALSE}
# loop over each bounding box, get the DSM raster 
# and find the AZ polygon and save as geoJSON

az_elev_m <- 25 # AZ is area -25m elevation from summit
unlink(paste0(this_dir, "/from_url_req/"),
       recursive = TRUE)

# n = 137
for(i in 133:nrow(gjsf_elev)){
  
  this_summit <- gjsf_elev[i, ] 
  this_square <- gjsf_elev_buf_sq_df[i, ]
  
  print(paste0("Starting work on the AZ for ", this_summit$summitCode,
               "..."))

# simplify bounding box coords
bbx_m <- 
  gjsf_elev_buf_sq_df[i, ] %>% 
  st_cast("POINT") %>% 
  st_coordinates()

# construct URL to query the LIDAR data portal using our bbox coords
url <- paste0("https://lidarportal.dnr.wa.gov/download?geojson=%7B%22type%22%3A%22Polygon%22%2C%22",
              "coordinates%22%3A%5B%5B%5B",
              bbx_m[1,1], "%2C", bbx_m[1,2], "%5D%2C%5B",
              bbx_m[2,1], "%2C", bbx_m[2,2], "%5D%2C%5B", 
              bbx_m[3,1], "%2C", bbx_m[3,2], "%5D%2C%5B",
              bbx_m[4,1], "%2C", bbx_m[4,2], "%5D%2C%5B",
              bbx_m[5,1], "%2C", bbx_m[5,2], "%5D%5D%5D%7D",
              "&ids=",
            "1615",   "%2C",    # East Cascades South 2020 DTM (not hillshade)
            "1609",   "%2C",     # East Cascades North 2020 DTM (not hillshade)
            "1501",    "%2C",      # King County East 2021 DTM (not hillshade)
            "1603",    "%2C",     # King County East 2021 DTM (not hillshade)
            "1231" ,    "%2C"  ,     # Yakima Basin North 2018
             "825",     "%2C",     # Colockum 2014
            "1463",    "%2C",     # Manastash 2018
             "733",     "%2C",    # Yakima K2K 2014
              "1555",      "%2C",     # Cle Elum Forest Studies 2021
              "40",     "%2C",     # Teanaway 2015
              "45",     "%2C",     # Wenatchee 2009
              "1265",    "%2C",     # Yakima Basin 2018
              "1495",    "%2C",     # Paco Basin 2020
              "877",     "%2C",     # Columbia 2010 B
            "1439",   "%2C" ,     # Cascades South 3DEP 2019
             "1525",  "%2C",     # Lewis East Sthelens 2020
             "1045",  "%2C",     # Mount Adams 2016
             "715",   "%2C",     # Wasco A 2015
             "1253",  "%2C",     # Southwest State Lands 2017
             "373",   "%2C",     # Wasco B 2015
             "36",    "%2C",     # Speelyai 2015
             "1075",  "%2C",     # Swwa Foothills 2017
             "889",   "%2C",     # Columbia 2010 E
             "1385",  "%2C",     # Southwest Wa Opsw 2019
             "19",    "%2C",      # Klickitat 2015 
             "1069",  "%2C",       # North Puget 2017
             "1118",  "%2C",       # North Cascades 2009
             "22",    "%2C",       # North Puget USGS 2006
             "43",    "%2C",       # Tulalip 2013
             "303",   "%2C",       # Glacier Peak 2015
             "810",   "%2C",       # Rainier 2007
             "24",   "%2C",        # Pierce 2011
             "777",   "%2C",      # San Juan 2009
             "15",   "%2C",       # Hood Canal 2015
             "1259",   "%2C",      # Kitsap County Opsw 
             "30",    "%2C",       # San Juan 2013 
             "16",    "%2C",       # Island 2014
             "1409",    "%2C",       # Entiat Upper 2018
             "1349",    "%2C",       # Chelan 2018
             "9",      "%2C",       # Entiat 2013 
             "667",    "%2C",       # Chelan 2015 
             "1173",    "%2C",       # Wenatchee Upper 2017
             "1451",    "%2C",       # Methow Basin 2018
             "703",    "%2C",       # Okanogan 2015
             "697",    "%2C",       # Loomis Loup Loup 2016
             "661",    "%2C",       # Baker 2015
             "755",    "%2C",       # Colville 2014
             "6",    "%2C",       # Colville 2008
             "679",    "%2C",       # Colville 2012
             "685",    "%2C",       # Colville 2015
             "1343",    "%2C",       # Ne Wa 2017
             "1051",    "%2C"       # Ne Wa 2016
) 

# when we get an error, search for the summit on sotl.as, open the summit in caltopo, paste the coords into the lidar URL, select the summit and see what LIDAR dataset is available. 

print(paste0("Downloading LIDAR data for ", this_summit$summitCode,
             "..."))

# send our query to the LIDAR portal, unzip the response and import the tif file
req <- request(url)
resp <- req %>% req_perform()
v = resp_body_raw(resp)
rm(resp)
writeBin(v, paste0(this_dir, "/data.zip"))
rm(v)
unzip(paste0(this_dir, "/data.zip"), 
      exdir = paste0(this_dir, "/from_url_req/"))
unlink(paste0(this_dir, "/data.zip"))
the_raster_files <-
  list.files(path = paste0(this_dir, "/from_url_req/"),
             pattern = "*.tif",
             full.names = TRUE,
             ignore.case = TRUE,
             recursive = TRUE)
#------------------------------------------------------------------------
# Manage the downloaded tif files, in case there are multiple with 
# different resolutions

# import into our R session, mostly there are multiple tif 
# files, but sometimes just one
if(length(the_raster_files) == 1){
  
  # just one raster
  m <- rast(the_raster_files)
  
} else {

  # more than one, and maybe with different resolutions
  
rasters_res_tbl <- 
tibble(
  file = the_raster_files,
  res = map_dbl(the_raster_files, ~res(rast(.x))[[1]])
) %>% 
  # make sure hi res is first
  arrange(desc(res))

res_n <- unique(rasters_res_tbl$res)

if(length(res_n) == 1 ) {
  # if all the same resolution 
     m <-  sprc(rasters_res_tbl$file[ rasters_res_tbl$res == res_n ])
     
           print(paste0("Merging LIDAR raster files with the same resolution for ", this_summit$summitCode,
             "..."))
           
     m <- terra::merge(m, gdal=c("BIGTIFF=YES", "NUM_THREADS = ALL_CPUS") )
} else {
  # if not all the same resolution 

# create a list of rasters where we've merged together files with the same
# resolution, if there are multiples
list_of_rasters <- vector("list", length = length(res_n))

for(i in 1:length(res_n)){
  
  the_res <- res_n[i]
  
  if(sum(rasters_res_tbl$res == the_res) == 1){
    # if just one raster of that resolution
    list_of_rasters[[i]] <- rast(rasters_res_tbl$file[ rasters_res_tbl$res == the_res ])
  } else {
    # if multiple raster files of that resolution
    
      print(paste0("Merging LIDAR raster files with the same resolution for ", this_summit$summitCode,
             "..."))
    
    list_of_rasters[[i]]  <-  sprc(rasters_res_tbl$file[ rasters_res_tbl$res == the_res ])
    list_of_rasters[[i]]  <- terra::merge(list_of_rasters[[i]], gdal=c("BIGTIFF=YES", "NUM_THREADS = ALL_CPUS") )
  }
  
}

# downsample the hi-res raster to match the low-res raster, assuming we
# have only two different resolutions of rasters here, and so only a list o
# of two items, hopefully this is true!

# in case CRS are not the same
  crs(list_of_rasters[[1]]) <-  crs(list_of_rasters[[2]])
  
  print(paste0("Merging LIDAR raster files with different resolutions for ", this_summit$summitCode,
             "..."))
  
  # https://gis.stackexchange.com/a/423700
  e <- exactextractr::exact_resample(list_of_rasters[[2]], #. low res raster,
                                     list_of_rasters[[1]], #  high res raster
                                     'mean')
  
  m <- terra::merge(e,   # output from previous
                    list_of_rasters[[1]])  # high res
  
}
}

#------------------------------------------------------------------------

# transform summit and bounding box coords 
# to the projection of LIDAR data
this_square_nad83 <- st_transform(this_square, st_crs(m))
this_summit_nad83 <- st_transform(this_summit, st_crs(m))

# subset LIDAR data that fills just this square
lidar_cropped <- 
  terra::crop(m, this_square_nad83)

# delete downloaded files
unlink(paste0(this_dir, "/from_url_req/"), recursive = TRUE)

# # take a look
# ggplot() +
#   geom_spatraster(data = lidar_cropped) +
#  #  geom_spatraster(data = m) +
#   geom_sf(data = this_summit_nad83) +
#   geom_sf(data = this_square_nad83,
#           fill = NA) +
#   scale_fill_viridis_c(na.value = "white",
#                        name = "Elevation (ft)") +
#   annotation_scale(location = "bl",
#                    width_hint = 0.5,
#                    pad_y = unit(0.1, "cm"),
#                    pad_x = unit(0.5, "cm"),
#                    style =  "ticks") +
#   coord_sf()
#------------------------------------------------------------------------

# crop the raster to the AZ
lidar_cropped_max_elev_ft <- minmax(lidar_cropped)[2]

# define elevation contour that bounds the AZ
this_summit_point_az <- 
  this_summit_nad83 %>% 
  mutate(lidar_cropped_max_elev_ft = lidar_cropped_max_elev_ft, 
         lidar_cropped_max_elev_m = lidar_cropped_max_elev_ft / 3.2808399,
         az_lower_contour = ifelse(elev_m <= lidar_cropped_max_elev_m,
                                   elev_m - az_elev_m,         
                                   lidar_cropped_max_elev_m - az_elev_m ),  # SOTA summit data does not always match raster data
         az_lower_contour_ft = az_lower_contour * 3.2808399) 

# subset the summit point that is in this bbox
lidar_cropped[lidar_cropped < this_summit_point_az$az_lower_contour_ft] <- NA

# # take a look
# ggplot() +
#   geom_spatraster(data = lidar_cropped) +
#   geom_sf(data = this_summit_nad83) +
#   geom_sf(data = this_square_nad83,
#           fill = NA) +
#   scale_fill_viridis_c(na.value = "white",
#                        name = "Elevation (ft)") +
#   annotation_scale(location = "bl",
#                    width_hint = 0.5,
#                    pad_y = unit(0.1, "cm"),
#                    pad_x = unit(0.5, "cm"),
#                    style =  "ticks") +
#   coord_sf()

#------------------------------------------------------------------------
# get extent of the AZ raster as polygon
az_poly <- st_as_sf(as.polygons(lidar_cropped > -Inf))

# ggplot() +
# geom_sf(data = this_summit_nad83) +
# geom_sf(data = az_poly,
#         colour = "red",
#         fill = NA) +
# scale_fill_viridis_c(na.value = "white",
#                      name = "Elevation (ft)") +
# annotation_scale(location = "bl",
#                  width_hint = 0.5,
#                  pad_y = unit(0.1, "cm"),
#                  pad_x = unit(0.5, "cm"),
#                  style =  "ticks") +
# coord_sf()

# if there are multiple polygons, we only want the one that 
# contains the summit point when we have multipolys, 
# we just want the one with the summit in it

# dissolve all into one polygon
df_union_cast <- sf::st_union(sf::st_as_sf(az_poly))
df_union_cast <- st_cast(df_union_cast, "POLYGON")

poly_with_summit <- 
  apply(st_is_within_distance(df_union_cast, 
                              this_summit_nad83, 
                              sparse = FALSE,
                              dist = 50), 2, # within or 10 m outside of, because some 
        # summits are just outside of their
        # nearest polygon
        function(col) { 
          df_union_cast[which(col), ]
        })[[1]]

# # now we see the single polygon that is the activation zone
# ggplot() +
#   geom_sf(data = poly_with_summit) +
#   geom_sf(data = this_summit_nad83) +
#   coord_sf() +
#   annotation_scale(location = "bl", 
#                    width_hint = 0.5,
#                    pad_y = unit(0.1, "cm"),
#                    pad_x = unit(0.5, "cm"),
#                    style =  "ticks")  
  

poly_with_summit <- st_as_sf(st_transform(poly_with_summit, st_crs(this_summit)))

#------------------------------------------------------------------------
# saving 

print(paste0("Saving GeoJSON file for ", this_summit$summitCode,
             "..."))

# write AZ polygon to a GeoJSON file
file_name <- paste0(paste0(this_dir, "/output/"), 
                    str_replace_all(this_summit$summitCode, "/|-", "_"), 
                    ".geojson")

# export AZ polygon as a GeoJSON file
geojsonio::geojson_write(poly_with_summit, 
                         file = here::here(file_name),
                         quiet = TRUE)


}

```

After running the code block above we now have one GeoJSON file per summit with a polygon defining its activation zone in our `/output` directory.

### Inspect all the summits and activation zones in our set

Import the GeoJSON files and display the summit points and activation zones polygons on an interactive map. 

```{r results='hide'}
library(tidyverse)
library(sf)

# import the GeoJSON files with the AZ polygons
gjson_files <- 
  list.files(path = paste0(this_dir, "/output"),
             full.names = TRUE,
             recursive = TRUE)

az_files <- 
  map(gjson_files,
      st_read,
      quiet = TRUE) 

names(az_files) <- str_remove_all(basename(gjson_files), 
                                  "output|.geojson")

summit_geometry <- 
  bind_rows(az_files, .id = "summit") %>% 
  select(summit, geometry)
```

This code block makes an interactive map of the summits and activation zones. 

```{r}
library(leaflet)

which_summit <- 
  which(summit_geometry$summit == "W7W_FR_136")

which_summit_lng_lat <- 
as.vector(st_coordinates(summit_geometry[which_summit, ])[1,1:2])

l <- 
leaflet() %>% 
  addProviderTiles("OpenStreetMap") %>% 
  addPolygons(data = summit_geometry,
              color = "red", 
              weight = 1) %>% 
  addCircleMarkers(data = gjsf %>% 
                     mutate(summitCode = str_replace_all(summitCode,
                                                         "/|-", "_")) %>% 
                     filter(summitCode %in% summit_geometry$summit[which_summit] ),
                   radius = 1,
                   weight = 1,
                   label = ~summitCode,
                   labelOptions = labelOptions(noHide = T, 
                                               direction = "top",
                                               offset = c(0, -15))) %>% 
  setView(lng = which_summit_lng_lat[1],
          lat = which_summit_lng_lat[2],  
           zoom = 16)  %>%
  addScaleBar()


l # open the interactive map for panning and zooming over all the summits
```



